---
title: "Dehumanization Project Data Processing"
output: html_notebook
---

The R Version used to create this file is: `R version 4.3.2 (2023-10-31)`. 
To reproduce analysis, switch with your own version of R by pressing `Tools` > `Global Options` > `R Version`


STEP 1. Installing groundhog and downloading packages. Not visible in output document.

```{r libraries setup, include = FALSE}
knitr::opts_knit$set(root.dir = '/Users/elizabethmiclau/Downloads/Dehumanization Project')


# Change out the working directory for your own


install.packages("remotes")
if(!suppressWarnings(require(groundhog))){remotes::install_github('CredibilityLab/groundhog')}
library("groundhog")

# put all packages in and call groundhog
pkgs <- c(
    "aws.s3",
    "tidyverse"
    )

pkgs_all <- c(pkgs)

#Change out the date for today's date (unless you are running an old analysis)
groundhog.library(pkgs_all, "2023-11-27")


```

STEP 2. Filtering files by size to figure out which data files are incomplete and move the incomplete ones into a newly created "did not finish" folder.

```{r downloading data and examining file sizes}
library(dplyr)
library(ggplot2)
library(purrr)
library(stringr)
library(readr)
library(fs)

# To run this code. you need to change the working directory (wd) for your computer. In this new wd, you need to have:
  # 1. A folder named 'data' with 54 participants' csv files (only keep the 54 files with long strings of letters/numbers0
  # 2. Downloaded survey data from Qualtrics, saved as a file named: surveydata.csv


setwd("~/Downloads/Dehumanization Project") # Confirming working directory to the folder containing data
data_folder <- '/Users/elizabethmiclau/Downloads/Dehumanization Project/data'
raw_files <- file.info(dir_ls(path = data_folder, regexp = "\\.csv$"))
raw_files$filename <- rownames(raw_files)  # Add a column with file paths

# Adding a column for file size in kilobytes
raw_files$size_kb <- raw_files$size / 1024

# Plotting file sizes to visually inspect for a cutoff point
ggplot(raw_files, aes(x = size_kb)) +
  geom_histogram(color = "black", bins = 30) +
  labs(title = "Histogram of File Sizes", x = "File Size (KB)", y = "Frequency")

# Define your file size criteria for incomplete (in KB)
size_criteria_kb <- 6.7

# Define and create the "did not finish" folder
incomplete_folder <- paste0(data_folder, "/did_not_finish")
dir.create(incomplete_folder, recursive = TRUE, showWarnings = FALSE)

# Find files smaller than the size criteria and move them
small_files <- raw_files$filename[raw_files$size_kb < size_criteria_kb]
for (file in small_files) {
  file.rename(from = file, to = file.path(incomplete_folder, basename(file)))
}

# Path to remaining CSV files
csv_files <- dir_ls(path = data_folder, regexp = "\\.csv$") 

# Combine the remaining CSV files into one data frame
experiment_data <- map_dfr(csv_files, ~read_csv(.x, show_col_types = FALSE))

# Creating a new data frame (CSV file) with experiment data
write.csv(experiment_data, "experiment_data.csv", row.names = FALSE)


```


STEP 3. Processing the experimental data - Wide long format
```{r cleaning the dataset}
experiment_data <- read.csv("~/Downloads/Dehumanization Project/experiment_data.csv")

# Viewing the data set
dim(experiment_data)   # dimensions
names(experiment_data) # variables
str(experiment_data)   # structure

# Removing and renaming columns
experiment_data <- experiment_data[ , !(names(experiment_data) %in% c('rt', 'view_history', 'internal_node_id','keep_remove','trial_type'))]
experiment_data <- experiment_data %>% 
                 rename(name = Name)
experiment_data <- experiment_data %>% 
                 rename(Prolific_ID = participant_id)
experiment_data <- experiment_data[c("Prolific_ID", "time_elapsed", "trial_index", "responses", 
                                 "stimulus", "button_pressed", "name", "trial", 
                                 "agent_names_trial", "agent_DH", "agent_NDH")]

# Convert units from 'time_elapsed' from ms to minutes and seconds
experiment_data$time_elapsed_min_sec <- sprintf("%d:%02d", 
                                              experiment_data$time_elapsed %/% 60000, 
                                              (experiment_data$time_elapsed %% 60000) %/% 1000)
experiment_data <- experiment_data[ , !(names(experiment_data) %in% c('time_elapsed'))]
experiment_data <- experiment_data %>%
  rename(time_elapsed = time_elapsed_min_sec)

# Ensuring each participant completed the full experiment (40-41 trial_index lines). 
participant_trial_count <- experiment_data %>%
  group_by(Prolific_ID) %>%
  summarise(trial_count = n_distinct(trial_index))
incomplete_participants <- participant_trial_count %>%
  filter(!(trial_count %in% c(40, 41)))
if (nrow(incomplete_participants) > 0) {
  print("Some participants did not complete the full experiment..")
  print(incomplete_participants)
} else {
  print("All participants completed the full experiment.")
}

# Checking participants ID
check_responses <- function(data) {
  data %>%
    filter(trial_index == 0) %>%
    mutate(check = str_detect(responses, paste0("\\{\"ID\":\"", Prolific_ID, "\"\\}"))) %>%
    filter(!check) %>%
    select(Prolific_ID, trial_index, responses)
}
discrepancies <- check_responses(experiment_data)
if (nrow(discrepancies) > 0) {
  print("Discrepancies found in the following rows:")
  print(discrepancies)
} else {
  print("All manually entered participant IDs match recorded IDs.")
}
experiment_data <- experiment_data %>%
  mutate(responses = ifelse(trial_index == 0, NA, responses))


# Checking that all participants passed the Attention Checks
# Attention Check 1: Giraffe
discrepancy_indices <- which(experiment_data$trial_index == 3 & 
                             !grepl("iraf|irraf|neck", experiment_data$responses))
if (length(discrepancy_indices) > 0) {
  cat("The following rows failed Attention Check #1")
  print(discrepancy_indices)
} else {
  cat("All participants pass Attention Check #1.")
}
experiment_data <- experiment_data %>%
  mutate(responses = ifelse(trial_index == 3, NA, responses))

# Attention Check 2: "but"
discrepancy_indices <- which(experiment_data$trial_index == 5 & 
                             !str_detect(experiment_data$responses, "but"))
if (length(discrepancy_indices) > 0) {
  cat("The following rows failed Attention Check #2.")
  print(discrepancy_indices)
} else {
  cat("All participants passed Attention Check #2.")
}
experiment_data <- experiment_data %>%
  mutate(responses = ifelse(trial_index == 5, NA, responses))

# Attention Check 3: "rock"
discrepancy_indices <- which(experiment_data$trial_index == 18 & 
                             !str_detect(experiment_data$responses, "rock"))
if (length(discrepancy_indices) > 0) {
  cat("The following rows failed Attention Check #3.")
  print(discrepancy_indices)
} else {
  cat("All participants passed Attention Check #3.")
}
experiment_data <- experiment_data %>%
  mutate(responses = ifelse(trial_index == 18, NA, responses))

# Cleaning the 'stimulus' column
experiment_data <- experiment_data %>%
  mutate(stimulus = ifelse(str_detect(stimulus, "<img src=\"img/stimuli/"), NA, stimulus))
experiment_data <- experiment_data %>%
  mutate(stimulus = gsub("img/stimuli/Slide([0-9]+)\\.png", "Slide \\1", stimulus))

# Cleaning the 'button_pressed' column
experiment_data <- experiment_data %>%
  mutate(button_pressed = ifelse(
    !str_detect(stimulus, "Slide") & 
    !str_detect(name, "gender|phone|computer"),
    NA, 
    button_pressed
  ))

experiment_data <- experiment_data %>%
  mutate(choice = sapply(1:nrow(.), function(i) {
    if(grepl("Slide", experiment_data$stimulus[i])) {
      # Split the agent names
      agents <- strsplit(as.character(experiment_data$agent_names_trial[i]), ",")[[1]]
      
      # Determine the chosen name based on button_pressed
      chosen_name <- ifelse(experiment_data$button_pressed[i] == 0, agents[1], agents[2])
      
      # Check which column (agent_DH or agent_NDH) matches the chosen name
      if(chosen_name == experiment_data$agent_DH[i]) {
        return("DH")
      } else if(chosen_name == experiment_data$agent_NDH[i]) {
        return("NDH")
      } else {
        return(NA)  # Return NA if no match is found
      }
    } else {
      return(NA)  # Return NA if stimulus does not contain "Slide"
    }
  }))


# Converting variables to correct format
experiment_data$stimulus <- factor(experiment_data$stimulus)
experiment_data$trial_index <- factor(experiment_data$trial_index)
experiment_data$button_pressed <- factor(experiment_data$button_pressed)
experiment_data$trial <- factor(experiment_data$trial)
experiment_data$agent_DH <- factor(experiment_data$agent_DH)
experiment_data$agent_NDH <- factor(experiment_data$agent_NDH)
experiment_data$choice <- factor(experiment_data$choice)
str(experiment_data) # Checking my work


# Cleaning up the environment
rm(discrepancies, incomplete_participants, participant_trial_count, small_files, size_criteria_kb, discrepancy_indices, file, raw_files)

# Reordering the columns
experiment_data <- experiment_data[c("Prolific_ID", "time_elapsed", "trial_index", "responses", 
                                 "stimulus", "button_pressed", "name", "trial", 
                                 "agent_names_trial", "agent_DH", "agent_NDH", "choice")]

experiment_data_widelong <- experiment_data
## EXPERIMENTAL DATA IN IS WIDE-LONG FORMAT ##
view(experiment_data_widelong)


# Creating a new data frame (CSV file) with experiment data
write.csv(experiment_data_widelong, "experiment_data_widelong.csv", row.names = FALSE)


```


STEP 4. Processing the experimental data - Long format
```{r Switching experiment_data to long format}

library(tidyr)
library(dplyr)
experiment_data_long <- read.csv("~/Downloads/Dehumanization Project/experiment_data_widelong.csv")

# Removing 'time_elapsed' and 'trial_index' columns
experiment_data_long <- experiment_data_long %>% 
  select(-trial_index)

# Creating 'gender' column
experiment_data_long <- experiment_data_long %>% 
  mutate(gender = case_when(
    name == "gender" & button_pressed == 0 ~ "Male",
    name == "gender" & button_pressed == 1 ~ "Female",
    TRUE ~ as.character(NA)
  ))

# Creating 'device' column
experiment_data_long <- experiment_data_long %>% 
  mutate(device = case_when(
    name == "phone" & button_pressed == 1 ~ "Phone",
    name == "computer" & button_pressed == 0 ~ "Computer",
    TRUE ~ as.character(NA)
  ))

# Removing all rows without a value in 'button_pressed' column
experiment_data_long <- experiment_data_long %>% 
  filter(!is.na(button_pressed))

# Checking for 12 rows per ppt
prolific_counts <- table(experiment_data_long$Prolific_ID)
if(any(prolific_counts != 12)) {
  warning("There are Prolific_IDs with counts not equal to 12")
}

# Creating left and right columns
experiment_data_long <- experiment_data_long %>% # Creating 'left_agent' and 'right_agent' columns
  mutate(left_agent = ifelse(grepl(",", agent_names_trial), sub(",.*", "", agent_names_trial), NA),
         right_agent = ifelse(grepl(",", agent_names_trial), sub(".*,", "", agent_names_trial), NA))
experiment_data_long <- experiment_data_long %>% # Creating 'left_rating' and 'right_rating' columns
  mutate(left_rating = ifelse(left_agent == agent_DH, "DH", "NDH"),
         right_rating = ifelse(right_agent == agent_DH, "DH", "NDH"))
experiment_data_long <- experiment_data_long %>% # Creating 'left_choice' and 'right_choice' columns
  mutate(left_choice = ifelse(choice == left_rating, 1, 0),
         right_choice = ifelse(choice == right_rating, 1, 0))

# Removing unneeded columns
experiment_data_long <- experiment_data_long %>% 
  select(-responses, -button_pressed, -name, -agent_names_trial, -agent_DH, -agent_NDH, -choice)

# Transforming the data from widelong to long format
experiment_data_long <- experiment_data_long %>%
  gather(key = "side", value = "agent", left_agent, right_agent) %>%
  gather(key = "rating_side", value = "rating", left_rating, right_rating) %>%
  gather(key = "choice_side", value = "choice", left_choice, right_choice) %>%
  filter(ifelse(side == "left_agent", rating_side == "left_rating", rating_side == "right_rating") &
         ifelse(side == "left_agent", choice_side == "left_choice", choice_side == "right_choice")) %>%
  select(-rating_side, -choice_side) %>%
  mutate(side = ifelse(side == "left_agent", "left", "right")) %>%
  arrange(Prolific_ID, trial, side)

# Remove rows where 'side' is "right" and 'gender' or 'device' is not NA
experiment_data_long <- experiment_data_long %>%
  filter(!(side == "right" & (!is.na(gender) | !is.na(device))))

# For rows where 'side' is "left" and 'gender' is not NA, shift 'gender' up and remove the row
rows_to_shift <- which(experiment_data_long$side == "left" & !is.na(experiment_data_long$gender))

# Creating a new data frame (CSV file) with experiment data
write.csv(experiment_data_long, "experiment_data_long.csv", row.names = FALSE)


```


STEP 5. Processing the survey data
```{r cleaning the survey data}
setwd("~/Downloads/Dehumanization Project") # Confirming working directory to the folder containing data
survey_data <- read.csv("~/Downloads/Dehumanization Project/surveydata.csv")
view(survey_data)

# Removing, renaming, and reordering columns
survey_data <- survey_data[ , !(names(survey_data) %in% c('Status', 'IPAddress', 'Progress','Finished','RecipientLastName','RecipientFirstName','RecipientEmail','ExternalReference','LocationLatitude','LocationLongitude','DistributionChannel','UserLanguage','StartDate','EndDate'))]
survey_data <- survey_data[-2, ]
rownames(survey_data) <- NULL
survey_data <- survey_data %>% 
                 rename(Duration_in_sec = Duration..in.seconds.,
                        Date = RecordedDate,
                        Prolific_ID = id,
                        Response_ID = ResponseId,
                        Image_Difficulties1 = Q13,
                        Image_Difficulties2 = Q16,
                        Comments_Questions = Q14)
survey_data <- survey_data[c("Response_ID", "Prolific_ID", "Date","Duration_in_sec","Q2_1", "Q3_1", 
                                 "Q3_2", "Q3_3", "Q3_4", "Q3_5", "Q3_6", "Q5_1", "Q5_2", "Q5_3", 
                                 "Q5_4", "Q5_5", "Q5_6","Q5_7", "Age","Gender", "Race", "Income", 
                                 "Conservatism", "Education", "Image_Difficulties1", 
                                 "Image_Difficulties2","Comments_Questions", "Data_Quality")]

# Replacing values with notes to just the number
survey_data[] <- lapply(survey_data, function(x) gsub("1 = Not at all", "1", x))
survey_data[] <- lapply(survey_data, function(x) gsub("4 = A moderate amount", "4", x))
survey_data[] <- lapply(survey_data, function(x) gsub("7 = An extreme amount", "7", x))
survey_data[] <- lapply(survey_data, function(x) gsub("1 = Strongly disagree", "1", x))
survey_data[] <- lapply(survey_data, function(x) gsub("7 = Strongly agree", "7", x))

# Converting all columns to the correct variable types
survey_data$Date <- as.Date(survey_data$Date, format = "%Y-%m-%d %H:%M:%S")
survey_data$Response_ID <- factor(survey_data$Response_ID) # Response_ID: chr to factor
survey_data$Response_ID[1] <- NA
survey_data$Prolific_ID <- factor(survey_data$Prolific_ID) # Prolific_ID: chr to factor
survey_data$Prolific_ID[1] <- NA
survey_data$Q3_1 <- factor(survey_data$Q3_1) # Q3_1: chr to factor
survey_data$Q3_2 <- factor(survey_data$Q3_2) # Q3_2: chr to factor
survey_data$Q3_3 <- factor(survey_data$Q3_3) # Q3_3: chr to factor
survey_data$Q3_4 <- factor(survey_data$Q3_4) # Q3_4: chr to factor
survey_data$Q3_5 <- factor(survey_data$Q3_5) # Q3_5: chr to factor
survey_data$Q3_6 <- factor(survey_data$Q3_6) # Q3_6: chr to factor
survey_data$Q5_1 <- factor(survey_data$Q5_1) # Q5_1: chr to factor
survey_data$Q5_2 <- factor(survey_data$Q5_2) # Q5_2: chr to factor
survey_data$Q5_3 <- factor(survey_data$Q5_3) # Q5_3: chr to factor
survey_data$Q5_4 <- factor(survey_data$Q5_4) # Q5_4: chr to factor
survey_data$Q5_5 <- factor(survey_data$Q5_5) # Q5_5: chr to factor
survey_data$Q5_6 <- factor(survey_data$Q5_6) # Q5_6: chr to factor
survey_data$Q5_7 <- factor(survey_data$Q5_7) # Q5_7: chr to factor
survey_data$Race <- factor(survey_data$Race) # Race: chr to factor
survey_data$Gender <- factor(survey_data$Gender) # Race: chr to factor
survey_data$Income <- factor(survey_data$Income) # Race: chr to factor
survey_data$Conservatism <- factor(survey_data$Conservatism) # Conservatism: chr to ordered factor
survey_data$Image_Difficulties1 <- factor(survey_data$Image_Difficulties1) # Image_Difficulties1: chr to factor
survey_data$Image_Difficulties2 <- factor(survey_data$Image_Difficulties2) # Image_Difficulties2: chr to factor
survey_data$Comments_Questions <- factor(survey_data$Comments_Questions) # Comments_Questions: chr to factor
survey_data$Data_Quality <- factor(survey_data$Data_Quality) # Data_Quality: chr to factor

survey_data$Conservatism <- factor(
  survey_data$Conservatism,
  levels = c("Extremely Liberal", "Liberal", "Somewhat Liberal", 
             "Middle of the Road", "Somewhat Conservative", 
             "Conservative", "Extremely Conservative"),
  ordered = TRUE
)
survey_data$Education <- factor( # Education: chr to ordered factor
  survey_data$Education,
  levels = c("No degree", "High school/GED", "Some college experience", 
             "2-year college degree (A.A.)", "4-year college degree (BA/BS)", 
             "Graduate or Professional Degree (MA/ MBA/ MS/ PhD)"),
  ordered = TRUE)

# Ignore Warnings for the following conversions (reason: First row cannot be a numerical value)
survey_data$Duration_in_sec <- as.numeric(survey_data$Duration_in_sec) # Duration: chr to num
survey_data$Q2_1 <- as.numeric(survey_data$Q2_1) # Q2_1: chr to num
survey_data$Age <- as.numeric(survey_data$Age) # Age: chr to num

# Reviewing responses to the Image Difficulty Question
survey_data$Image_Difficulties2[survey_data$Image_Difficulties2 == ""] <- NA
print(levels(survey_data$Image_Difficulties2))

# Reviewing responses to the Image Difficulty Question
survey_data$Comments_Questions[survey_data$Comments_Questions == ""] <- NA
survey_data$Comments_Questions <- as.character(survey_data$Comments_Questions)
survey_data$Comments_Questions[survey_data$Comments_Questions %in% c("No", "None", "no", "No comments", "I have no comments. ")] <- NA
survey_data$Comments_Questions <- factor(survey_data$Comments_Questions)
print(levels(survey_data$Comments_Questions))
  
view(survey_data)


```


STEP 6. Merging Survey Data with Experimental Data - Wide-long format 
```{r integrating survey_data into experiment_data_widelong}

# Convert 'Prolific_ID' to character in both datasets
experiment_data_widelong$Prolific_ID <- as.character(experiment_data_widelong$Prolific_ID)
survey_data$Prolific_ID <- as.character(survey_data$Prolific_ID)


# Check if 'Prolific_ID' exists in both datasets and has the same data type
if("Prolific_ID" %in% names(experiment_data_widelong) & "Prolific_ID" %in% names(survey_data)) {
  if(class(experiment_data_widelong$Prolific_ID) == class(survey_data$Prolific_ID)) {

    full_data_widelong <- merge(experiment_data_widelong, survey_data, by = "Prolific_ID", all.x = TRUE, all.y = TRUE)
    
  } else {
    stop("Data types of 'Prolific_ID' in both datasets do not match.")
  }
} else {
  stop("'Prolific_ID' column not found in one or both datasets.")
}

view(full_data_widelong)

# Creating a new data frame (CSV file) with merged experimental and survey data
write.csv(full_data_widelong, "full_data_widelong.csv", row.names = FALSE)

```

