---
title: "Dehumanization Project Data Processing"
output: html_notebook
---

The R Version used to create this file is: `R version 4.3.2 (2023-10-31)`. 
To reproduce analysis, switch with your own version of R by pressing `Tools` > `Global Options` > `R Version`


STEP 1. Installing groundhog and downloading packages. (Not visible in the output document.)

```{r libraries setup, include = FALSE}
knitr::opts_knit$set(root.dir = '/Users/elizabethmiclau/Downloads/Dehumanization Project')


install.packages("remotes")
if(!suppressWarnings(require(groundhog))){remotes::install_github('CredibilityLab/groundhog')}
library("groundhog")


# put all packages in and call groundhog
pkgs <- c(
    "aws.s3",
    "tidyverse"
    )

pkgs_all <- c(pkgs)

#Change out the date for today's date (unless you are running an old analysis)
groundhog.library(pkgs_all, "2023-11-27")


```



STEP 2. Filtering files by size to figure out which data files are incomplete and move the incomplete ones into a newly created "did not finish" folder.

```{r downloading data and examining file sizes}

library(dplyr)
library(ggplot2)
library(purrr)
library(stringr)
library(readr)
library(fs)

setwd("~/Downloads/Dehumanization Project") # Set the working directory to the folder containing data
data_folder <- "data"

raw_files <- file.info(dir_ls(path = data_folder, regexp = "\\.csv$"))
raw_files$filename <- rownames(raw_files)  # Add a column with file paths

# Adding a column for file size in kilobytes
raw_files$size_kb <- raw_files$size / 1024

# Plotting file sizes to visually inspect for a cutoff point
ggplot(raw_files, aes(x = size_kb)) +
  geom_histogram(color = "black", bins = 30) +
  labs(title = "Histogram of File Sizes", x = "File Size (KB)", y = "Frequency")

# Define your file size criteria for incomplete (in KB)
size_criteria_kb <- 6.7

# Define and create the "did not finish" folder
incomplete_folder <- paste0(data_folder, "/did_not_finish")
dir.create(incomplete_folder, recursive = TRUE, showWarnings = FALSE)

# Find files smaller than the size criteria and move them
small_files <- raw_files$filename[raw_files$size_kb < size_criteria_kb]
for (file in small_files) {
  file.rename(from = file, to = file.path(incomplete_folder, basename(file)))
}

# Path to remaining CSV files
csv_files <- dir_ls(path = data_folder, regexp = "\\.csv$") 

# Combine the remaining CSV files into one data frame
combined_data <- map_dfr(csv_files, read_csv)

# Optional: Write the combined data frame to a new CSV file
write.csv(combined_data, "combined_data.csv", row.names = FALSE)


```


STEP 3. Process the data in the next few chunks. The final output of this data processing should be one long-format csv that contains one trial per line.
```{r cleaning the dataset}

combined_data <- read.csv("~/Downloads/Dehumanization Project/combined_data.csv")

# Viewing the data set
dim(combined_data)   # dimensions
names(combined_data) # variables
str(combined_data)   # structure

# Removing and reordering columns
combined_data <- combined_data[ , !(names(combined_data) %in% c('rt', 'view_history', 'internal_node_id','keep_remove','trial_type'))]
combined_data <- combined_data[c("participant_id", "trial_index", "responses", 
                                 "stimulus", "button_pressed", "Name", "trial", 
                                 "agent_names_trial", "agent_DH", "agent_NDH", "time_elapsed")]

# Convert units from 'time_elapsed' from ms to minutes and seconds
combined_data$time_elapsed_min_sec <- sprintf("%d:%02d", 
                                              combined_data$time_elapsed %/% 60000, 
                                              (combined_data$time_elapsed %% 60000) %/% 1000)
combined_data <- combined_data[ , !(names(combined_data) %in% c('time_elapsed'))]
combined_data <- combined_data %>%
  rename(time_elapsed = time_elapsed_min_sec)

# Ensuring each participant completed the full experiment (40-41 trial_index lines). 
participant_trial_count <- combined_data %>%
  group_by(participant_id) %>%
  summarise(trial_count = n_distinct(trial_index))
incomplete_participants <- participant_trial_count %>%
  filter(!(trial_count %in% c(40, 41)))
if (nrow(incomplete_participants) > 0) {
  print("Some participants did not complete the full experiment..")
  print(incomplete_participants)
} else {
  print("All participants completed the full experiment.")
}

# Checking participants ID
check_responses <- function(data) {
  data %>%
    filter(trial_index == 0) %>%
    mutate(check = str_detect(responses, paste0("\\{\"ID\":\"", participant_id, "\"\\}"))) %>%
    filter(!check) %>%
    select(participant_id, trial_index, responses)
}
discrepancies <- check_responses(combined_data)
if (nrow(discrepancies) > 0) {
  print("Discrepancies found in the following rows:")
  print(discrepancies)
} else {
  print("All manually entered participant IDs match recorded IDs.")
}
combined_data <- combined_data %>%
  mutate(responses = ifelse(trial_index == 0, NA, responses))


# Checking that all participants passed the Attention Checks
# Attention Check 1: Giraffe
discrepancy_indices <- which(combined_data$trial_index == 3 & 
                             !grepl("iraf|irraf|neck", combined_data$responses))
if (length(discrepancy_indices) > 0) {
  cat("The following rows failed Attention Check #1")
  print(discrepancy_indices)
} else {
  cat("All participants pass Attention Check #1.")
}
combined_data <- combined_data %>%
  mutate(responses = ifelse(trial_index == 3, NA, responses))

# Attention Check 2: "but"
discrepancy_indices <- which(combined_data$trial_index == 5 & 
                             !str_detect(combined_data$responses, "but"))
if (length(discrepancy_indices) > 0) {
  cat("The following rows failed Attention Check #2.")
  print(discrepancy_indices)
} else {
  cat("All participants passed Attention Check #2.")
}
combined_data <- combined_data %>%
  mutate(responses = ifelse(trial_index == 5, NA, responses))

# Attention Check 3: "rock"
discrepancy_indices <- which(combined_data$trial_index == 18 & 
                             !str_detect(combined_data$responses, "rock"))
if (length(discrepancy_indices) > 0) {
  cat("The following rows failed Attention Check #3.")
  print(discrepancy_indices)
} else {
  cat("All participants passed Attention Check #3.")
}
combined_data <- combined_data %>%
  mutate(responses = ifelse(trial_index == 18, NA, responses))

# Cleaning the 'stimulus' column
combined_data <- combined_data %>%
  mutate(stimulus = ifelse(str_detect(stimulus, "<img src=\"img/stimuli/"), NA, stimulus))
combined_data <- combined_data %>%
  mutate(stimulus = gsub("img/stimuli/Slide([0-9]+)\\.png", "Slide \\1", stimulus))

# Cleaning the 'button_pressed' column
combined_data <- combined_data %>%
  mutate(button_pressed = ifelse(
    !str_detect(stimulus, "Slide") & 
    !str_detect(Name, "gender|phone|computer"),
    NA, 
    button_pressed
  ))


# Viewing newly formatted data set
view(combined_data)

# Converting variables to correct format
str(combined_data)

# Convert time_elapsed from chr to time string
install.packages("lubridate") # package for time conversion
library(lubridate)
combined_data$time_elapsed <- as.hms(hms(combined_data$time_elapsed))
## THIS IS NOT WORKING YET








```

